Unsupervised embeddings capture latent structure of scientific mobility

### 2020-02-24 paper/current/main.tex
First results para about correlation findings

\subsection*{Data}
%% Main dataset
% Do we need to add the Leiden field classificaiton infomration? It doesn't seem like its relevant for the current analysis.
% Do we need to talk about how we define mobility in relation to past papers?
We source co-affiliation trajectories of authors from the Web of Science database hosted by the Center for Science and Technology Studies at Leiden University. 
Trajectories are constructed from author affiliations listed on the byline of publications for a given author.
Given the limitations of author-name disambiguation, we limit to papers published after 2008, when the Web of Science began providing full names and institutional affiliations~\cite{caron2014disambiguation} and when performance is strongest (ref to SI?). 
This yields 33,934,672 author-affiliation combinations representing 112,963,792 authors. 
Each author-affiliation combination is associated with the publication year and a unique id that maps the affiliation to a disambiguate organization. 
Combinations are also associated with a publication-level disciplinary classification, one of ``Social sciences and humanities'',
``Biomedical and health sciences'', ``Physical sciences and engineering'', ``Life and earth sciences'', and ``Mathematics and computer science''.
Trajectories are represented as the list of author-affiliation combinations, ordered by year of publication, and randomly ordered for combinations within the same year. 
Authors are classified as mobile when they have at least two distinct organization ids in their trajectory, meaning that they have published using two or more distinct affiliations between 2008 and 2019.

@@

%
% SUBSECTION: SEMAXIS
%
\subsection{SemAxis}

@@

>>>>

\subsection*{Data}
%% Main dataset
% Do we need to add the Leiden field classificaiton infomration? It doesn't seem like its relevant for the current analysis.
% Do we need to talk about how we define mobility in relation to past papers?
We source co-affiliation trajectories of authors from the Web of Science database hosted by the Center for Science and Technology Studies at Leiden University. 
Trajectories are constructed from author affiliations listed on the byline of publications for a given author.
Given the limitations of author-name disambiguation, we limit to papers published after 2008, when the Web of Science began providing full names and institutional affiliations~\cite{caron2014disambiguation} and when performance is strongest (ref to SI?). 
This yields 33,934,672 author-affiliation combinations representing 112,963,792 authors. 
Each author-affiliation combination is associated with the publication year and a unique id that maps the affiliation to a disambiguate organization. 
Combinations are also associated with a publication-level disciplinary classification, one of ``Social sciences and humanities'',
``Biomedical and health sciences'', ``Physical sciences and engineering'', ``Life and earth sciences'', and ``Mathematics and computer science''.
Trajectories are represented as the list of author-affiliation combinations, ordered by year of publication, and randomly ordered for combinations within the same year. 
Authors are classified as mobile when they have at least two distinct organization ids in their trajectory, meaning that they have published using two or more distinct affiliations between 2008 and 2019.
While the number of publications has increased year-to-year, the mobility and disciplinary makeup of the dataset has not significantly changed across the period of study (see Fig.~\ref{fig:supp:pubs_over_time}).

@@

@@

We find that embedding distance, the distance between vectors in the embedding space, produces an effective measure of distance.
Embedding distance explains more of the total flux between pairs of organizations, with correlation ($R^{2} = 0.46$, Fig.~\ref{fig:gravity_performance}a) over double that of geographic distance ($R^{2} = 0.22$, Fig.~\ref{fig:gravity_performance}b).
Similarly, embedding distance outperforms personalized-page rank distance ($R^{2} = 0.001$, Fig. SI~\ref{fig:supp:gravity_ppr}a), a measure produced from the factorized nodes in the underlying mobility network. 
When controlling for only mobility between pairs of organizations within the same country, embedding distance ($R^{2} = 0.48$, Fig. SI~\ref{fig:supp:gravity_correlation}c) than does geographic distance ($R^{2} = 0.19$, Fig. SI~\ref{fig:supp:gravity_correlation}d). 
Embedding distnace is also far more robust in explaining international flux, with a correlation nearly seven times higher ($R^{2} = 0.28$, Fig. SI~\ref{fig:supp:gravity_correlation}e) than that of geographic distance ($R^{2} = 0.041$, Fig. SI~\ref{fig:supp:gravity_correlation}f).
